Plots

### 2025-05-08 manuscript/manuscript.tex
Replaced main text figures, working on editing text

@@

Of the 6939 records, GPT-4 had the highest PCCC followed by GPT-3.5, InSilicoVA, and InterVA-5 at 0.61, 0.58, 0.44, and 0.43 PCCC respectively (Table \ref{tab-perf-overall}). GPT-3.5 and GPT-4 had improvements ranging from 0.14-0.18 PCCC over InSilicoVA and InterVA-5, while GPT-4 slightly improved over GPT-3.5 by 0.03 PCCC. CSMF accuracies were above 0.7 for all models at 0.79, 0.74, 0.74, and 0.78 for GPT-4, GPT-3.5, InSilicoVA, and InterVA-5 respectively. Figure \ref{fig-perf-agegroup} shows the PCCC performance across age groups with reference to the CSMF Accuracy. GPT-4 had the highest PCCC at 0.65 and 0.62 for the adult and neonatal age groups, while GPT-3.5 had the highest PCCC at 0.57 for the child age group with GPT-4 performing slightly worse at 0.54. GPT-3.5 had the second highest PCCC at 0.56 fo the adult age group, while InSilicoVA had the second highest PCCC for the neonatal age group at 0.56. InSilicoVA and InterVA-5 performed the worse for both adult and child age groups at PCCC scores below 0.5, while GPT-3.5 performed the worse for the neonatal age group with a PCCC of 0.42. The PCCC for models had larger differences between the minimum and maximum PCCC for the adult age group at 0.24, than the child and neonatal age groups at 0.12 and 0.20 respectively. For reference, all records, including non-physician agreed records, were also coded by GPT-3.5, GPT-4, InterVA-5, and InSilicoVA for comparison to physician agreed records. For physician agreed records, models had 0.05 to 0.09 higher PCCC values and 0.01 to 0.05 higher CSMF accuracy values than for all records (see Figure \ref{fig-perf-allvsagree} in Appendix \ref{app-perf-misc}).

\begin{table}[h]
\caption{Overall model performance on CGHR-10 CODs for physician agreed records.}\label{tab-perf-overall}%
\begin{tabular}{@{}llll@{}}
\toprule
Model & PCCC & CSMF Accuracy & PCCC Improvement\\
\midrule
GPT-4 & 0.61 & 0.79 & $\uparrow$ 0.03\\
GPT-3.5 & 0.58 & 0.74 & $\uparrow$ 0.14\\
InSilicoVA & 0.44 & 0.74 & $\uparrow$ 0.01\\
InterVA-5 & 0.43 & 0.78 & -\\
\botrule
\end{tabular}
\end{table}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{fig-perf-agegroup.pdf}
\caption{Model performance by age group.}\label{fig-perf-agegroup}
\end{figure}

@@

Across the adult age ranges, GPT-4 had the highest PCCC values at or above 0.59, followed by GPT-3.5, while InterVA-5 and InSilicoVA had the lowest PCCC values at or below 0.46. GPT-4 PCCC was between 0.61 and 0.72 for young to middle age ranges between 10 to 49 years, and at or below 0.62 for older age ranges between 50 to 69 years (Figure \ref{fig-perf-agerange-adult}). Each age range had between 135 (4\%) to 575 (15\%) records with the 65-69 year age range having the most records (n=575, 15\%), while other age ranges ranged from 4-9\% of all records. Figure \ref{fig-perf-cod-adult} shows the highest and lowest performing models by PCCC across 17 adult CODs excluding suicide due to the less notable 3 (\textless 1\%) records. The highest PCCC values were GPT-4 between 0.36 to 0.99 for 10 of 17 CODs (GPT-3.5, InterVA-5, and InSilicoVA within 0.05 for 3, 1, and 1 of these 10 CODs respectively), InSilicoVA for tuberculosis and road and transport injuries at 0.72 and 0.84 (GPT-4 and InterVA-5 within 0.05), and GPT-3.5 between 0.44 and 0.94 for 5 of 17 CODs (GPT-4 within 0.05 for malaria). The lowest PCCC values were InterVA-5 between 0 to 0.79 for 8 of 17 CODs (InSilicoVA and GPT-3.5 within 0.05 for cancers and unspecified infections respectively), InSilicoVA between 0.01 to 0.41 for 6 of 17 CODs (InterVA-5 within 0.05 for other cardiovascular diseases and diarrhoeal diseases), and GPT-3.5 at 0.39 for tuberculosis and 0.54 for road and transport injuries. Chronic respiratory diseases had the largest differences between minimum and maximum PCCC at 0.93. Maternal conditions, and road and transport injuries, had higher PCCC values above or at 0.67 (with the exception of GPT-3.5 for road and transportation injuries as an outlier), where there were small differences between minimum and maximum PCCC below or at 0.2. Ischemic heart disease and unspecified infections had PCCC values below or at 0.65, where there were small differences between minimum and maximum PCCC at or below 0.14. Each adult COD had between 27 (\textless 1\%) to 799 records (21\%) with malaria (n=799, 21\%), other noncommunicable diseases (n=414, 11\%), and unspecified infections (n=389, 10\%) making 42\% of all records, and other CODs between 1-9\% of all records. GPT-4 had the highest PCCC for both male and female records at 0.65 and 0.64 respectively, while all other models had PCCC at or below 0.6 (see Figure \ref{fig-perf-sex-adult} in Appendix \ref{app-perf-misc}).

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-perf-cod-adult.pdf}
\caption{Model performance for adult records by COD.}\label{fig-perf-cod-adult}
\end{figure}

@@

Across the child age ranges, the highest PCCC was observed for GPT-4 at 0.59 and 0.66 for 1-5 years and 6-11 years (GPT-3.5 within 0.05), GPT-3.5 at 0.54 for 6-11 months, and InSilicoVA at 0.5 for 1-5 months (Figure \ref{fig-perf-agerange-child}). Each child age range had between 308 (12\%) to 1633 (62\%) records with most records in the 1-5 years range at 1633 records (62\%) and the other age ranges consisting of 12-14\% of all records. For the 8 child CODs (excluding congenital anomalies due to small sample size of 1 (\textless 1\%) record), the highest PCCC observed was for GPT-4 for 4 of 8 CODs, GPT-3.5 for 3 of 8 CODs (InSilicoVA within 0.05 for other infections), and InSilicoVA for pneumonia (Figure \ref{fig-perf-cod-child}). The lowest PCCC across the 8 child CODs was observed for InSilicoVA for epilepsy, leukaemia, and other noncommunicable diseases and malaria, InterVA-5 for pneumonia, nutritional deficiencies, and other infections. GPT-3.5, InSilicoVA, and InterVA-5 all had similar PCCC values greater than 0.8 as the lowest PCCC for injuries. Epilepsy, leukaemia, and other noncommunicable disease had the largest difference between minimum and maximum PCCC at 0.68, while Ill-defined had the next largest difference at 0.67 with InSilicoVA and InterVA-5 at 0 PCCC. Malaria and other infections had low PCCC values below 0.58 across models with small differences between minimum and maximum PCCC at 0.18 and 0.14 respectively. Injuries and diarrhoeal diseases had high PCC values above 0.6 across models with small differences between minimum and maximum PCC at 0.14 and 0.3 respectively. Each of the 8 child CODs had between 11 (\textless 1\%) to 1382 (52\%) records with malaria (n=1382, 52\%), other infections (n=667, 25\%), and diarrhoeal diseases (n=175, 7\%) making 84\% of all records. Nutritional deficiencies had less than 1\% of all records, while other CODs had between 1-3\% of all records. GPT-3.5 had the highest PCCC for both male and female records at 0.56 and 0.58 with GPT-4 within 0.05 PCCC at 0.55 and 0.54 respectively, while all other models had PCCC below 0.5 (see Figure \ref{fig-perf-sex-child} in Appendix \ref{app-perf-misc}).

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{fig-perf-cod-child.pdf}
\caption{Model performance for child records by COD.}\label{fig-perf-cod-child}
\end{figure}

@@

GPT-4 had the highest PCCC for 0-6 day and 7-27 day neonatal age ranges at 0.6 and 0.7 respectively, while GPT-3.5 had the lowest PCCC across all neonatal age ranges as seen in Figure \ref{fig-perf-agerange-neo}. There were large differences between minimum and maximum PCCC for 7-27 days at 0.33, and small differences for 0-6 days at 0.16. Most records were in the 0-6 day age range at 395 (83\%) records, while the rest were in the 7-27 day age range at 82 (17\%) records. Across 5 neonatal CODs (excluding congenital anomalies and other due to small sizes of 2 (\textless 1\%) and 5 (1\%) respectively), the highest PCCC was observed for GPT-4 for 3 of the 5 CODs (GPT-3.5 identical for prematurity and low birthweight), GPT-3.5 for 2 of the 5 CODs (InSilicoVA identical for stillbirth), and InSilicoVA for neonatal infections (Figure \ref{fig-perf-cod-neo}). Stillbirth had small differences between minimum and maximum PCCC at 0.07. The other 4 CODs had large differences between minimum and maximum PCCC ranging from 0.46 to 0.64. Each neonatal COD had between 23 (5\%) to 172 (38\%) records, with stillbirth (n=172, 36\%), birth asphyxia and birth trauma (n=103, 22\%), and neonatal infections (n=99, 21\%), and prematurity and low birthweight (n=73, 15\%) making 94\% of all records, while the rest are in ill-defined. GPT-4 had the highest PCCC for both female and male records at 0.62, while all other models had PCCC at or below 0.56 (see Figure \ref{fig-perf-sex-neo} in Appendix \ref{app-perf-misc}).

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{fig-perf-cod-neo.pdf}
\caption{Model performance for neonatal records by COD.}\label{fig-perf-cod-neo}
\end{figure}

@@

PJ and PB are the study Principal Investigators. ATA and RK implemented the data collection procedures. RW, CC, LN, and TKSN processed, documented, and prepared the data. RW, ASL, and RK ran the models. RW wrote the paper and conducted the analysis. AB provided medical domain guidance and feedback. All authors reviewed the results and contributed to the report. All authors read and approved the final manuscript.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-data-sampling.pdf}
\caption{Case study data sampling areas.}\label{fig-data-sampling}
\end{figure}

@@

\section{Performance By Other Variables}\label{app-perf-misc}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-perf-allvsagree.pdf}
\caption{Model performance for all records versus physician agreed records.}\label{fig-perf-allvsagree}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-perf-agerange-adult.pdf}
\caption{Model performance for adult records by five-year age range.}\label{fig-perf-agerange-adult}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-perf-sex-adult.pdf}
\caption{Model performance for adult records by sex.}\label{fig-perf-sex-adult}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{fig-perf-agerange-child.pdf}
\caption{Model performance for child records by five-year age range.}\label{fig-perf-agerange-child}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-perf-sex-child.pdf}
\caption{Model performance for child records by sex for assigning CGHR-10 CODs.}\label{fig-perf-sex-child}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{fig-perf-agerange-neo.pdf}
\caption{Model performance for neonatal records by age range.}\label{fig-perf-agerange-neo}
\end{figure}

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-perf-sex-neo.pdf}
\caption{Model performance for neonatal records by sex for assigning CGHR-10 CODs.}\label{fig-perf-sex-neo}
\end{figure}

>>>>

\author[1]{\fnm{Ronald} \sur{Carshon-Marsh}}\email{ronald.carshonmarsh@mail.utoronto.ca}

@@

Of the 6939 records, GPT-4 had the highest PCCC followed by GPT-3.5, InSilicoVA, and InterVA-5 at 0.61, 0.56, 0.44, and 0.43 PCCC respectively (Figure \ref{fig-pccc-csmf}). GPT-3.5 and GPT-4 had improvements ranging from 0.14-0.18 PCCC over InSilicoVA and InterVA-5, while GPT-4 slightly improved over GPT-3.5 by 0.03 PCCC. CSMF accuracies were above 0.7 for all models at 0.79, 0.74, 0.74, and 0.78 for GPT-4, GPT-3.5, InSilicoVA, and InterVA-5 respectively. Figure \ref{fig-pccc} shows the PCCC performance across age groups with reference to the CSMF Accuracy. GPT-4 had the highest PCCC at 0.65 and 0.62 for the adult and neonatal age groups, while GPT-3.5 had the highest PCCC at 0.57 for the child age group with GPT-4 performing slightly worse at 0.54. GPT-3.5 had the second highest PCCC at 0.56 fo the adult age group, while InSilicoVA had the second highest PCCC for the neonatal age group at 0.56. InSilicoVA and InterVA-5 performed the worse for both adult and child age groups at PCCC scores below 0.5, while GPT-3.5 performed the worse for the neonatal age group with a PCCC of 0.42. The PCCC for models had larger differences between the minimum and maximum PCCC for the adult age group at 0.24, than the child and neonatal age groups at 0.12 and 0.20 respectively. For reference, all records, including non-physician agreed records, were also coded by GPT-3.5, GPT-4, InterVA-5, and InSilicoVA for comparison to physician agreed records. For physician agreed records, models had 0.05 to 0.09 higher PCCC values and 0.01 to 0.05 higher CSMF accuracy values than for all records.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{fig-pccc-csmf.pdf}
    \caption{Overall model performance.}\label{fig-pccc-csmf}
    \end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{fig-pccc.pdf}
\caption{Model performance by age group.}\label{fig-pccc}
\end{figure}

@@

Figure \ref{fig-pccc-cod-adult} shows the highest and lowest performing models by PCCC across 17 adult CODs excluding suicide due to the less notable 3 (\textless 1\%) records. The highest PCCC values were GPT-4 between 0.36 to 0.99 for 10 of 17 CODs (GPT-3.5, InterVA-5, and InSilicoVA within 0.05 for 3, 1, and 1 of these 10 CODs respectively), InSilicoVA for tuberculosis and road and transport injuries at 0.72 and 0.84 (GPT-4 and InterVA-5 within 0.05), and GPT-3.5 between 0.44 and 0.94 for 5 of 17 CODs (GPT-4 within 0.05 for malaria). The lowest PCCC values were InterVA-5 between 0 to 0.79 for 8 of 17 CODs (InSilicoVA and GPT-3.5 within 0.05 for cancers and unspecified infections respectively), InSilicoVA between 0.01 to 0.41 for 6 of 17 CODs (InterVA-5 within 0.05 for other cardiovascular diseases and diarrhoeal diseases), and GPT-3.5 at 0.39 for tuberculosis and 0.54 for road and transport injuries. Chronic respiratory diseases had the largest differences between minimum and maximum PCCC at 0.93. Maternal conditions, and road and transport injuries, had higher PCCC values above or at 0.67 (with the exception of GPT-3.5 for road and transportation injuries as an outlier), where there were small differences between minimum and maximum PCCC below or at 0.2. Ischemic heart disease and unspecified infections had PCCC values below or at 0.65, where there were small differences between minimum and maximum PCCC at or below 0.14. Each adult COD had between 27 (\textless 1\%) to 799 records (21\%) with malaria (n=799, 21\%), other noncommunicable diseases (n=414, 11\%), and unspecified infections (n=389, 10\%) making 42\% of all records, and other CODs between 1-9\% of all records.

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-pccc-cod-adult.pdf}
\caption{Model performance for adult records by COD.}\label{fig-pccc-cod-adult}
\end{figure}

@@

For the 8 child CODs (excluding congenital anomalies due to small sample size of 1 (\textless 1\%) record), the highest PCCC observed was for GPT-4 for 4 of 8 CODs, GPT-3.5 for 3 of 8 CODs (InSilicoVA within 0.05 for other infections), and InSilicoVA for pneumonia (Figure \ref{fig-pccc-cod-child}). The lowest PCCC across the 8 child CODs was observed for InSilicoVA for epilepsy, leukaemia, and other noncommunicable diseases and malaria, InterVA-5 for pneumonia, nutritional deficiencies, and other infections. GPT-3.5, InSilicoVA, and InterVA-5 all had similar PCCC values greater than 0.8 as the lowest PCCC for injuries. Epilepsy, leukaemia, and other noncommunicable disease had the largest difference between minimum and maximum PCCC at 0.68, while Ill-defined had the next largest difference at 0.67 with InSilicoVA and InterVA-5 at 0 PCCC. Malaria and other infections had low PCCC values below 0.58 across models with small differences between minimum and maximum PCCC at 0.18 and 0.14 respectively. Injuries and diarrhoeal diseases had high PCC values above 0.6 across models with small differences between minimum and maximum PCC at 0.14 and 0.3 respectively. Each of the 8 child CODs had between 11 (\textless 1\%) to 1382 (52\%) records with malaria (n=1382, 52\%), other infections (n=667, 25\%), and diarrhoeal diseases (n=175, 7\%) making 84\% of all records. Nutritional deficiencies had less than 1\% of all records, while other CODs had between 1-3\% of all records.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{fig-pccc-cod-child.pdf}
\caption{Model performance for child records by COD.}\label{fig-pccc-cod-child}
\end{figure}

@@

Across 5 neonatal CODs (excluding congenital anomalies and other due to small sizes of 2 (\textless 1\%) and 5 (1\%) respectively), the highest PCCC was observed for GPT-4 for 3 of the 5 CODs (GPT-3.5 identical for prematurity and low birthweight), GPT-3.5 for 2 of the 5 CODs (InSilicoVA identical for stillbirth), and InSilicoVA for neonatal infections (Figure \ref{fig-pccc-cod-neo}). Stillbirth had small differences between minimum and maximum PCCC at 0.07. The other 4 CODs had large differences between minimum and maximum PCCC ranging from 0.46 to 0.64. Each neonatal COD had between 23 (5\%) to 172 (38\%) records, with stillbirth (n=172, 36\%), birth asphyxia and birth trauma (n=103, 22\%), and neonatal infections (n=99, 21\%), and prematurity and low birthweight (n=73, 15\%) making 94\% of all records, while the rest are in ill-defined.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{fig-pccc-cod-neo.pdf}
\caption{Model performance for neonatal records by COD.}\label{fig-pccc-cod-neo}
\end{figure}

@@

PJ and PB are the study Principal Investigators. ATA and RK implemented the data collection procedures. RW, and TKSN processed, documented, and prepared the data. RW, ASL, and RK ran the models. RW wrote the paper and conducted the analysis. AB and RCM provided medical domain guidance and feedback. All authors reviewed the results and contributed to the report. All authors read and approved the final manuscript.

@@

\begin{figure}[H]
\centering
\includegraphics[width=\textwidth]{fig-map.pdf}
\caption{Case study data sampling areas.}\label{fig-map}
\end{figure}

@@

### 2025-05-08 src/plots.Rmd
Replaced main text figures, working on editing text

    # Get custom colors for models
    default_colors <- scales::hue_pal()(length(unique(df_plot$model_name)))
    names(default_colors) <- unique(df_plot$model_name)
    default_colors["Mean"] <- "black"

    # Plot trend lines for each model
    out <- ggplot(
        df_plot,
        aes(
            x = .data[[group]],
            y = metric,
            color = model_name,
            group = model_name
        )
    ) +
        geom_line(linewidth = 0.75, alpha = 0.4) + 
        geom_line(
            data = df_plot %>%
                group_by(.data[[group]]) %>%
                summarize(
                    metric = mean(metric),
                    model_name = "Mean"
                ),
            aes(color = model_name),
            se = F,
            linewidth = 1.25
        ) +
        scale_color_manual(values = default_colors) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            panel.border = element_rect(
                color = "black",
                fill = NA,
                linewidth = 0.75
            ),
            legend.title = element_blank(),
            legend.text = element_text(size = legend_text_size),
            legend.position = legend_position, 
            legend.direction = legend_direction,
            legend.box.spacing = legend_spacing,
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )
    return(out)

>>>>

    # Get custom colors for models
    default_colors <- scales::hue_pal()(length(unique(df_plot$model_name)))
    names(default_colors) <- unique(df_plot$model_name)
    default_colors["Mean Across Models"] <- "black"

    # Plot trend lines for each model
    out <- ggplot(
        df_plot,
        aes(
            x = .data[[group]],
            y = metric,
            color = model_name,
            group = model_name
        )
    ) +
        geom_line(linewidth = 0.75, alpha = 0.4) + 
        geom_line(
            data = df_plot %>%
                group_by(.data[[group]]) %>%
                summarize(
                    metric = mean(metric),
                    model_name = "Mean Across Models"
                ),
            aes(color = model_name),
            se = F,
            linewidth = 1.25
        ) +
        scale_color_manual(values = default_colors) +
        labs(
            title = title,
            x = x_title,
            y = y_title
        ) +
        theme_minimal() +
        theme(
            panel.border = element_rect(
                color = "black",
                fill = NA,
                linewidth = 0.75
            ),
            legend.title = element_blank(),
            legend.text = element_text(size = legend_text_size),
            legend.position = legend_position, 
            legend.direction = legend_direction,
            legend.box.spacing = legend_spacing,
            plot.title = element_text(size = title_size),
            axis.title.x = element_text(size = x_title_size),
            axis.title.y = element_text(size = y_title_size),
            axis.text.x = element_text(size = x_text_size),
            axis.text.y = element_text(size = y_text_size)
        )
    return(out)
